---
title: " "
output: 
    github_document
---

### https://github.com/hauselin/Rcode

#### Source or load `summaryh()` function

```{r, message=F, warning=F}
source("https://raw.githubusercontent.com/hauselin/Rcode/master/summaryh.R")
```

#### Linear regression standard R output 

```{r}
model <- lm(mpg ~ drat, mtcars)
summary(model)
```

#### Linear regression `summaryh()` output

```{r}
summaryh(model)
```

#### Linear regression `summaryh()` output: changing default output

```{r}
summaryh(model, decimal = 1)
```

#### Linear regression `summaryh()` output: changing default output

```{r}
summaryh(model, showTable = T)
```

#### Linear regression `summaryh()` output: show other effect sizes

```{r}
summaryh(model, showEffectSizesTable = T)
```

#### Correlation 

```{r}
summaryh(cor.test(mtcars$mpg, mtcars$cyl))
```

#### t-test

```{r}
summaryh(t.test(mpg ~ vs, mtcars))
```

#### ANOVA

```{r}
summaryh(aov(mpg ~ gear, mtcars), decimal = 1)
```

#### Multi-level/mixed effects model with lme4 and lmerTest

```{r, message=F,warning=F}
library(lme4); library(lmerTest) # load packages to fit mixed effects models
model <- lmer(weight ~ Time * Diet  + (1 + Time | Chick), data = ChickWeight)
summaryh(model)
```

### https://github.com/hauselin/Rcode
### https://github.com/hauselin/RDataScience
