---
title: "Tutorial 6: Simulating data"
author: Hause Lin
output: 
    github_document:
        toc: true
        # toc_float: false
        # theme: lumen
        # highlight: tango
editor_options: 
  chunk_output_type: console
---

## Loading frequently-used packages with `library()`

I always load my frequently-used packages at the top of each script.

```{r loading packages, results="hide", message=FALSE, warning=FALSE}
library(tidyverse); library(data.table); library(broom); library(dtplyr); library(lme4); library(lmerTest); library(ggbeeswarm); library(cowplot)
```

## Generate random values from a normal distribution

```{r}
rnorm(n = 10) # 10 values (values from standard normal distribution: mean 0, sd 1)
```

```{r}
rnorm(n = 100, mean = 100, sd = 15) # 100 values from a normal distribution (mean 100, sd = 15)
```

```{r}
x <- rnorm(n = 1000, mean = 100, sd = 15) # 1000 values from a normal distribution (mean 100, sd = 15)
hist(x) # base R plot histogram
```

## Generate random values from a uniform distribution

```{r}
x <- runif(n = 1000, min = 10, max = 20)
hist(x)
```

## Simulate a t-test with `rnorm()`

Group a (n = 20): mean 100, sd 15
Group b (n = 20): mean 115, sd 20

```{r}
fakeData <- data_frame(group = rep(c("a", "b"), each = 20), 
                       value = c(rnorm(20, 100, 15), rnorm(20, 115, 20))) 
print(fakeData, n = Inf)
```

Will group differences be significant?

```{r}
simulateResults <- t.test(value ~ group, data = fakeData)
simulateResults
source("https://raw.githubusercontent.com/hauselin/Rcode/master/summaryh.R")
summaryh(simulateResults) # my custom function
```

## Manipulating strings

Substitute patterns

```{r}
gsub(pattern = "a", replacement = "_HEY_", x = c("aba", "cae", "xxx"))
gsub(pattern = "a", replacement = " ", x = c("aba", "cae", "xxx"))
gsub(pattern = "a", replacement = "", x = c("aba", "cae", "xxx"))
```

Find matching patterns

```{r}
text <- c("foot","lefroo", "bafOobar", "zeb")
grep(pattern = "oo", x = text, ignore.case = T)
grep(pattern = "oo", x = text, ignore.case = F)
grepl(pattern = "oo", x = text, ignore.case = T)
grepl(pattern = "oo", x = text, ignore.case = F)
```

### Manipulating string resources

* [stringr tutorial/vignette](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html)
* [stringr github site](https://github.com/tidyverse/stringr)

## the apply family in R (looping without for loops)

* `lapply`, `apply`, `sapply`, `vapply`, `tapply`
