---
title: "Tutorial 2: tidyverse and datatable"
author: Hause Lin
output: 
    github_document:
        toc: true
        # toc_float: false
        # theme: lumen
        # highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r loading packages, results="hide", message=FALSE, warning=FALSE}
library(tidyverse); library(data.table); library(broom); library(dtplyr); library(lme4); library(lmerTest); library(ggbeeswarm); library(cowplot)
```

## Reading data into R

Read file in a directory and save the data as an object in the environment by using the assignment `<-` operator. 

```{r read data, results="hide", message=FALSE, warning=FALSE}
df1 <- read.csv("./Data/sleep.csv") # base R read.csv() function
# same as df1 <- read.csv("Data/sleep.csv")
# READ: assign the output read.csv("Data/sleep.csv") into df1

df2 <- fread("./Data/sleep.csv") # fread() from library(data.table)
# same as df2 <- fread("Data/sleep.csv")

# my favorite way of reading data: combining tbl_dt() and fread()
df3 <- tbl_dt(fread("./Data/sleep.csv")) # tbl_dt() from library(dtplyr)
# same as df3 <- tbl_dt(fread("Data/sleep.csv"))
```

The `.` in the file path simply refers to the current working directory, so it can be dropped. And `..` can be used to refer to the parent directory.

## Comparing the outputs of read.csv(x), fread(x), and tbl_dt(fread(x))

```{r}
df1 # read.csv("./Data/sleep.csv")
class(df1) # read.csv("./Data/sleep.csv")
```

```{r}
df2 # fread("./Data/sleep.csv") # fread() from library(data.table)
class(df2) # fread("./Data/sleep.csv") # fread() from library(data.table)
```

How's the output different from the one above?

```{r}
df3 # tbl_dt(fread("./Data/sleep.csv")) # tbl_dt() from library(dtplyr)
class(df3) # tbl_dt(fread("./Data/sleep.csv")) # tbl_dt() from library(dtplyr)
```

How's the output different from the two outputs above?

### Reading URLs and other formats

Check out the csv (comma separated values) data [here](https://raw.githubusercontent.com/hauselin/RDataScience/master/Data/sleep.csv). You can read data directly off a website.

Most of these read functions can import/read different types of files (e.g., csv, txt, URLs) as long as the raw data are formatted properly (e.g., separated by commas, tabs). But if you're trying to read proprietary data formats (e.g., SPSS datasets, Excel sheets), you'll need to use other libraries (e.g., `readxl`, `foreign`) to read those data into R.

```{r}
df_url <- tbl_dt(fread("https://raw.githubusercontent.com/hauselin/RDataScience/master/Data/sleep.csv"))
df_url # print data to console; same dataset as tbl_dt(fread("./Data/sleep.csv"))
```

## Summarizing objects

You can summarize objects quickly by using `summary()`, `str()`, `glimpse()`, or `print(x, n)`.

To view the first/last few items of an object, use `head()` or `tail()`.

```{r}
summary(df1) # we use summary() for many many other purposes
str(df1) 
glimpse(df1)
head(df1)
head(df3, n = 3) # what does this do?
tail(df3, n = 2)
print(df3, n = Inf) # Inf is infinte (so print all rows)
```

Use pipes `%>%` to summarize objects

```{r}
df3 %>% head(n = 2)
df3 %>% head(2) # why does this work?
df3 %>% print(Inf)
df3 %>% summary() # does this work? why?
```

Datatables and `dplyr`'s data_frame are much better because they tell us the class of each variable at the top when you print the object in the console. They also tell you the dimensions of your data.

```{r}
df3 # data_frame (dplyr) + data.table + tibble
```

## Using `$` and `[]` to extract elements using their names

```{r} 
names(df3)
df3$extra # extracts column/variable as a vector
df3$group
df3$ID
```

```{r}
myList <- list(a = -999, b = c(TRUE, FALSE, T, T), c = c('myAmazingList')) # create a list with named items a, b, c
class(myList)
names(myList) 
myList # note the formatting/structure of a list ($ signs tell you how to extract items)
myList$a
myList$b # lists can store objects of different classes (unlike most other objects, which expect all items to be of same class)
myList$c
```

```{r}
df1['extra'] # same as df$extra, but uses characters (in '') to extract elements
```

** BUT the syntax above only works for the `data.frame` class!**

```{r, error=TRUE}
df3['extra'] # fails!!!
```

If it's a `data.table` class, you do it differently (so know the classes of your objects)! More on `data.table` later on.

```{r}
df3[, extra] # df3[i, j] (i is row, and j is column)
```

## tidyverse: a collection of R packages

[tidyverse:](https://www.tidyverse.org/packages/)

>  The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures. 

Included packages: `ggplot2`, `dplyr`, `tidyr`, `stringr` etc. [see official website for documentation](https://www.tidyverse.org/packages/)

```{r}
library(tidyverse)
```

## Supercharging your workflow with data.table()

```{r}
df3[, extra]
df3[, .(extra)] # what's the difference? compare the classes of df3[, extra] and df3[, .(extra)]
```

