---
title: "Tutorial 5: Cleaning and analyzing questionnaires/surveys (reliability)"
author: Hause Lin
output: 
    github_document:
        toc: true
        # toc_float: false
        # theme: lumen
        # highlight: tango
editor_options: 
  chunk_output_type: console
---

## Loading frequently-used packages with `library()`

I always load my frequently-used packages at the top of each script.

```{r loading packages, results="hide", message=FALSE, warning=FALSE}
rm(list = ls()) # clear environment
library(tidyverse); library(data.table); library(broom); library(dtplyr); library(lme4); library(lmerTest); library(ggbeeswarm); library(cowplot)
```

## Read data (Qualtrics survey data) 

Here's a raw dataset collected using Qualtrics

```{r}
df1 <- tbl_dt(fread("./Data/qualtricsSurvey.csv"))
df1
glimpse(df1) # something looks weird and wrong!
```

With Qualtrics datasets, the first row (or even first few rows) of the dataset might contain irrelevant variables and information. If they contain extra rows, do the following to get rid of rows (commented out for now). But play with your dataset to see which rows you should remove.

In this example, row 1 contains extra data (the actual questions themselves), so we're removing it. 

```{r}
df1 <- tbl_dt(fread("./Data/qualtricsSurvey.csv", header = T))[-1] # delete row 1
# df1 <- tbl_dt(fread("./Data/qualtricsSurvey.csv"))[-c(1:5)] # delete rows 1:5 (example)
df1
glimpse(df1)
names(df1)
```

Cleaning data. If my id variable is all in numbers, I often make sure they are numerics!

```{r}
df1$ID # not numeric
```

Convert to numerics by applying the `as.numeric()` function

```{r}
df1[, ID := as.numeric(ID)] 
```

## Select questionnaire items for relability analysis

* use `select()` and `starts_with`, `matches()`, `contains()`, and `ends_with()` to select questionanire items for a scale

```{r}
scale_implicitBeliefs <- select(df1, starts_with("ImplicitB"))
scale_implicitBeliefs
```

Note that the variable types are all characters. Use `sapply()` (a loop) to loop through each column/variable, and convert them to `numeric` class. 

```{r}
scale_implicitBeliefs$ImplicitB_1 # all those quotation marks tell you it's a character
```

```{r}
scale_implicitBeliefsNumeric <- sapply(scale_implicitBeliefs, as.numeric) # convert to numeric (also a matrix)
scale_implicitBeliefsNumeric
```

I usually prefer to convert column classes using `data.table` syntax using `.SD` (*S*ubset of *D*atatable). Below, it's just saying, for every column in the `data.table`, convert it to numeric by applying the `as.numeric()` function to every column.

```{r}
scale_implicitBeliefsNumeric <- scale_implicitBeliefs[, lapply(.SD, as.numeric)]
scale_implicitBeliefsNumeric
```

Run reliability analysis

```{r}
library(psych) # psych package has the alpha() function to compute reliability
```

```{r}
scale_implicitBeliefsReliability <- alpha(scale_implicitBeliefsNumeric)
scale_implicitBeliefsReliability
```

Very low reliabilty! The function warns you if it thinks certain items should be reverse-scored!

You can set `check.keys = T` to automatically reverse-score items!

```{r}
scale_implicitBeliefsReliability <- alpha(scale_implicitBeliefsNumeric, check.keys = T) # automatically reverse-scores items if necessary
scale_implicitBeliefsReliability
```

Now reliability is much higher!

## Use `cleanQuestionnaire()` function instead

What if you want to compute each subject's scale mean? You can use `data.table` and `tidyverse` to do it. But with many scales, it can get tedious. Hence, I've written the `cleanQuestionnaire()` to automate certain steps. To use the function, first run `source('https://raw.githubusercontent.com/hauselin/Rcode/master/cleanQuestionnaire.R')` to get the function. Check out the actual [website itself](https://raw.githubusercontent.com/hauselin/Rcode/master/cleanQuestionnaire.R) if you wanto read the documentation of the function.

```{r}
source("https://raw.githubusercontent.com/hauselin/Rcode/master/cleanQuestionnaire.R")
```

Select data. Make sure the first column is subject id, and subsequent columns are items for a specific scale

* select id column (required)
* select items of a particular scale

```{r}
scale_implicitBeliefs <- select(df1, ID, starts_with("ImplicitB"))
scale_implicitBeliefs
```

Convert all columns to numeric

```{r}
scale_implicitBeliefs <- scale_implicitBeliefs[, lapply(.SD, as.numeric)]
scale_implicitBeliefs
```


Run `cleanQuestionnaire()` function

The function will double-check with you regarding which items to reverse-code, and which subscale an item belongs to. If everything looks correct, press 1 and hit enter to continue. 

The function accepts characters AND numerics scale items!

More documentation to follow [here](https://github.com/hauselin/Rcode) in the near future...

Function arguments

* data: data containing subject id and scale items (wide form)
* subjectCol: indicate which column contains subject id (default = 1)
* scaleName: give your scale a good name
* scaleMin: what's the scale mininum
* scaleMax: what's the scale maximum
* subscales: use a list() to specify item numbers for each subscale (if no subscale, just don't specify this subscales argument)
* itemsToReverse: item numbers to reverse-code
* checkReliability: T or F (whether to do reliability analysis) (does it separately for each subscale)

```{r, eval=F}
scale_implicitBeliefs_clean <- cleanQuestionnaire(data = scale_implicitBeliefs, subjectCol = 1, scaleName = "implicitBeliefs", scaleMin = 1, scaleMax = 6, subscales = list(implicitBeliefsA = c(1, 3, 5), implicitBeliefsB = c(2, 4, 6)), itemsToReverse = c(1, 2, 5), checkReliability = T)
```

```{r, echo=F}
scale_implicitBeliefs_clean <- read_rds("./Extra/scaleInfo.rds")
```

The output is a `list` class.

```{r}
class(scale_implicitBeliefs_clean)
```

```{r}
names(scale_implicitBeliefs_clean)
```

It contains the wide data, long data, and reliability results. To subset lists, use `$` to extract by name.

Get wide data: id variable, scale overall mean, and all subscale means

```{r}
scale_implicitBeliefs_clean$wide
```

Get long data: id variable, scale/subscale means (_m), standard deviations (_stdev), and range (_rge)

```{r}
scale_implicitBeliefs_clean$long
```

Get reliablity results (one set of results for each subscale)

```{r}
scale_implicitBeliefs_clean$reliability
```

## Another `cleanQuestionnaire()` example

Select data. Make sure the first column is subject id, and subsequent columns are items for a specific scale

* select id column (required)
* select items of a particular scale

```{r}
scale_workload <- select(df1, ID, starts_with("Workload"))
scale_workload
```

Run function

* no subscales, 1 item to reverse

```{r, eval=F}
scale_workload_clean <- cleanQuestionnaire(data = scale_workload, subjectCol = 1, scaleName = "workload", scaleMin = 1, scaleMax = 21, itemsToReverse = c(4), checkReliability = T)
```

```{r, echo=F}
scale_workload_clean <- read_rds("./Extra/scaleInfo2.rds")
```

Check results

```{r}
scale_workload_clean$wide
```

```{r}
scale_workload_clean$long
```

```{r}
scale_workload_clean$reliability
```

## Joining and saving wide data from multiple scales

Use `left_join()` to join two scales (both are wide form)

```{r}
joined <- left_join(scale_implicitBeliefs_clean$wide, scale_workload_clean$wide, by = 'id')
joined
```

## Save joined data as .rds object (one object at a time)

You can run the line below (without the # sign) to save any object as a .rds file. 

```{r, eval=F}
write_rds(joined, "scale_subject_means.rds") # save to your current directory
```

To read that back into R later on, use `read_rds()`

```{r, eval=F}
joined <- read_rds("scale_subject_means.rds") # read .rds file in your current directory into R
```

## Save your entire workspace and all objects in environment as .RData (all objects)

You can run the line below (without the # sign) to save any object as a .RData file. 

```{r, eval=F}
save.image("analysis_scales.RData") # save to your current directory
```

To read that back into R later on, use `load()`

```{r, eval=F}
load("analysis_scales.Rdata") # read .rds file in your current directory into R
```



