---
title: "Intro to data analysis with data.table package"
output: html_notebook
    toc: true
---

# data.table magic

Forget about <u>objects</u>, <u>variables</u>, <u>functions</u>, and all that programming jargon. Let's dive straight into data analysis! Before we begin each R session, you need to set up your <u>environment</u> (I'll try to underline all technical terms/jargon the first time I use them). 

```{r set up environment}
rm(list = ls()) # clear environment
setwd("/Users/Hause/Dropbox/Working Projects/RDataScience/R Basics/") # set working directory
library(data.table) # load packages
```

* `rm(list = ls())`: `rm()` is a function to **r**e**m**ove objects in your environment; `ls()` is a function to <u>print</u> the objects in your environment; so if you put `ls()` inside `rm()`, you're basically <u>calling</u> two functions (nested functions) this way: first get me all the objects in the environment (via `ls()`), then remove all of these objects (via `rm()`), hence `rm(list = ls())`
* `setwd("/Users/Hause/Dropbox/Working Projects/RDataScience/R Basics/")`: sets your <u>current working directory</u>

Ctrl+Shift+H

for

R has many built-in datasets. We'll use a few of them along the way. To see what's available, type `data()` in your Console to see them. 

```{r}
?sleep # 
dt <- data.table(sleep) 
class(dt)
dt
```


## data.table [i, j, by, .SDcols] 

`data.table` is an extremely flexible and powerful library/package. I think it should be in every data scientist's toolkit; and anyone who analyses data in R should know both `data.table` and `dplyr` (part of `tidyverse`) really well.

`data.table` uses very concise syntax (i.e., the way to express yourself) to perform powerful computations really quickly and flexibly. First, you need to understand its syntax: `data.table [i, j, by, .SDcols]`

* `i`: <u>filter</u> or <u>subset</u> rows/observations
* `j`: <u>select</u> particular columns/variables and apply any conceivable function to that column (or columns) 
* `by`: where all the magic happens; applies functions (specified in `j`) separately to different groups in your data
* `.SDcols`: applies functions (specified in `j`) to only a **s**ubset of the **d**atatable **col**umns

```{r}
dt


```

# t.test

```{r}
model_ttest <- t.test(formula = extra ~ group, data = dt)
model_ttest
```

```{r}
source("https://raw.githubusercontent.com/hauselin/Rcode/master/summaryh.R")
summaryh(model_ttest)
```

```{r}
model_lm <- lm(extra ~ group, dt)
summary(model_lm)
summaryh(model_lm)
```

```{r}
model_ttestOneSample <- t.test(dt[group == "1", extra], mu = 0)
model_ttestOneSample
summaryh(model_ttestOneSample)
```

